### 2017-01-22 notes.Rmd

# head direction cells

[@Taube:2007fk]

- mental simulation cannot just er-use the state used in sensory inference, because the velocity is fed to the head directino cells. The brain would get confused

- directinal signal is not just a simple integration of velocity signals from the vestibular system
- signals from motor efferecence copy or proprioceptive systems appear to be crucial for upating the preferred firing direction over time
- what about this? you don't know where you are, because you were passively moved. Can you update if I tell you (verbally instruct you)?
- vestibular input is eesntial for establishing the head direction signal? How can we reconcile this with the assumption that re-afference is cancelled, i.e. input due to active motion is cancelled. Maybe it's the control model...

# Corollary discharge
[@Crapse:2008jc]

# Predictive control 
[@Stahl:2011dl]

- movemements can induce sensory input
- sensory input due to own actions is indistinguishable from input due to external forces
- strategy: efference copy -> corollary discharge

it seems intuitively obvious that this mechanism has something to do with mental imagery.
This has been pointed out many times; mental imagery and mental simulation more generelly is somehow related to the capacity to compute corollay discharge, based on the efference copy and a forward model.

> Necessary components:  efference copy and a forward model

Sensory receptors don't care about their input. The brain has to deal with the problem further downstream