### 2017-01-27 notes.Rmd

---
output: 
  html_document: 
    fig_caption: yes
    number_sections: yes
    theme: spacelab
---

# one
that the brain is able to imagine, plan, perform counterfactual simulations because it coopts the mechanism evolved in order to predict the sensory consequences of its actions.

interference arises when he brain, due to task demands, uses the control mechanism in a manner which is not entirely counterfactual

# two
interference is surprising, because a mental simulation should be counterfactual. if it isn't this might be a problem. why should thinking about rotating left have an effect on your ability to sense leftward motion?

we show this type of interactions needs to be investigate pdf in a probabilistic framework

# three
we identify 2 main problem that have not been addressed: in SPT how do people decide which way and for how long to rotate?

and how does imagery interfere?

the first issue speaks to the fact that we have to solve the inverse problem before we can simulate in order to transform our FOR. the is equivalent to solving a planning problem, I.e. choosing the best action in order to reach a goal. this is actually counterfactual inference.

after introducing the graphical model, we show the types of inference, I.e. which nodes are clamped.

then we explain interference as biased perceptual decision making, concentrating on the Nigmatullina effect. how do biases arise? the basic model is not equipped to deal with bias.

# four
but how does interference through CVS on mental spatial transformation work? We hypothesize that CVS is a strange stimulus and needs interpretation by the brain. Passive actual movement puts the brain in an alert mode, because it might have to counteract the motion, I.e. the control circuits are engaged, leading to less capacity for planning.

in any case, we need a closer look at the connection between menral simulation and planning and the ability to predict self initiated movements.

# five
finally, how can we explain the bias in the Nigmatullina experiment? we need to investigate whether it is due to a bias in the starting point, in terms of a traditional method of analyzing such data, i.e. DDM. this is implemented by a prior on the choice alternatives, leading to starting point or altered drift rate.

or is it a model misspecification?

The alternative model by Merfeld would say it's just the starting point.

# six
perform simulations to show a theoretical analysis of what may cause the Nigmatullina effect.

model misspecification might be unlikely, but certainly worth a look

# seven
we tell the story in two parts.

1) describe how we can simulate using a parsimonious model that refuses components of the sensory inference model and components involved in distinguishing active from passive movements.

2) what happens when we simulate. switch off inputs, sensory gating corresponding to not resampling. show tables of correspondence mechanisms. show that interest,S of probabilistic computations the ability to model sensory consequences is a question of creating a proposal distribution for sampling from the prior. show table comparing Monterey carlo simulation with particle filtering. first show tables explaining components of filtering.

# eight
of course we first elucidate how the ability to simulate is related to sensory inference in real time.

vestibular sensory inference is a very good example, because it se ms to be well described as sequential monte carlo

# nine
interference is either:

1) mis-specified proposal distribution
2) biased decision-making

# 10
we focus explicitly know:

- forward simulation, explaining the requirements from a statistical / probabilistic point of view
- interference from thought to inference, and explore how this might occur. we focus on this direction because it is easier to understand than the other direction

show a sequence diagram showing a typical trial in an experiment

# eleven
we implement forward sampling by sampling from the proposal distribution without updating. this corresponds to monte carlo simulation

In order to set up a particular simulation, the brain must create a proposal distribution such that it represents the kinematics of the system to be modelled.